Vagueness, Noise, and Communication

### 2014-08-26 data_exploration_tmp.Rmd
data exploration MF

``` {r}
data = read.csv('results/20140813-194306.csv', 
                colClasses=c('numeric','factor','numeric','numeric','numeric','factor','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','character','character'))

@@

pairs(~ Number.of.states + Impairment + Speaker.entropy + Hearer.entropy +
        Speaker.convexity + Hearer.convexity + Speaker.Voronoiness + Hearer.Voronoiness +
        Speaker.informativity + Hearer.informativity + Expected.utility + Iterations, data=ds)
```

@@

pairs(~ Impairment + Speaker.entropy +
        Speaker.informativity + Hearer.informativity + Expected.utility, data=ds)
```

@@

also speaker and hearer voronoiness seems to be much higher for at least some bit of impairment

>>>>

``` {r}

require('ellipse')
require('ggplot2')

data = read.csv('results/20140813-194306.csv', 
                colClasses=c('numeric','factor','numeric','numeric','numeric','factor','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','character','character'))

@@

pairs(~ Number.of.states + Impairment + Speaker.entropy + Hearer.entropy +
        Speaker.convexity + Hearer.convexity +
        Speaker.informativity + Hearer.informativity + Expected.utility + Iterations, data=ds)
```

@@

dss = ds[,c(2,4,10,11,12)]

ctab <- cor(dss)

plotcorr(ctab)

# pairs(~ Impairment + Speaker.entropy +
#         Speaker.informativity + Hearer.informativity + Expected.utility, data=dss)
```

@@

also speaker and hearer voronoiness seems to be much higher for at least some bit of impairment, but we left this notion out [why what that again?]

Faster learning by impairment
-----------------------------

if learning is faster for some impairment, we should see these effects also in terms of higher EU at earlier stages

- plot EU at different stages in the nRD
- show that early on slight impairment is beneficial to no impairment

Next up
-------
- evaluate impact of number of states
- replot Fig 1 with seperated numbers of states

``` {r}

x = data

ggplot(x, aes(x=factor(Impairment), y=Expected.utility, color=factor(Number.of.states))) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(EU))

```

### 2014-08-26 paper/02_noisyRMD.tex
data exploration MF

Let the sender's (receiver's) behavioral strategies be functions in
$\Sstrat \in \Delta(\Messgs)^\States$ ($\Rstrat \in
\Delta(\Acts)^\Messgs$). Every mixed strategy $\Smixed$ converts to a
unique behavioral strategy defined by:
\begin{align*}
  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
\end{align*} 
Let $G$ be this mapping from mixed to behavioral strategies. Notice
that $G$ is \emph{not} an injection, as many mixed strategies map onto
the same behavioral strategy.

@@

We look at signaling games with $\States = \Acts$ and fix a confusion
matrix $C$, which is a row-stochastic matrix whose elements $C_{ij}$
give the probability that $\state_i$ is perceived as
$\state_j$. Define players average utility at each choice point given
the opponent's strategy as:
\begin{align*}
  \Phi(\state,\Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
\EU(\messg,\state,\Rstrat) &
\Phi(\messg,\Sstrat) & = \sum_{\state} \Rstrat(\state \probbar \messg)
\cdot \EU(\state,\messg,\Sstrat)\,.
\end{align*}
The discrete noise-perturbed replicator dynamics on behavioral
strategies proposed by \citet{Correia2013:The-Bivalent-Tr} is:
\begin{align*}
  \Sstrat'(\messg \probbar \state_i) & = \sum_{j} C_{ji} \cdot
  \frac{\Sstrat(\messg \probbar \state_j) \cdot
    \EU(\messg,\state_j,\Rstrat)} {\Phi(\state_j,\Rstrat)} \\
    \Rstrat'(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
  \frac{\Rstrat(\state_j \probbar \messg) \cdot
    \EU(\state_j,\messg,\Sstrat)} {\Phi(\messg,\Sstrat)}  \,.
\end{align*}
Notice that this update is also sequential: first we calculate an
update according to the standard replicator dynamics (in behavioral
strategies), then we compute the perturbation from perceptual noise.

>>>>

Let the sender's (receiver's) behavioral strategies be functions
$\Sstrat \in \Delta(\Messgs)^\States$ ($\Rstrat \in
\Delta(\Acts)^\Messgs$). Every mixed strategy $\Smixed$ converts to a
unique behavioral strategy defined by:
\begin{align*}
  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
\end{align*} 
Let $G$ be this mapping from mixed to behavioral strategies. Notice
that $G$ is \emph{not} an injection, as many mixed strategies map onto
the same behavioral strategy.

@@

We look at signaling games with $\States = \Acts$ and fix a confusion
matrix $C$, which is a row-stochastic matrix whose elements $C_{ij}$
give the probability that $\state_i$ is perceived as
$\state_j$. Define the players' average utility at each choice point given
the opponent's strategy as:
\begin{align*}
  \Phi(\state,\Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
\EU(\messg,\state,\Rstrat) &
\Phi(\messg,\Sstrat) & = \sum_{\state} \Rstrat(\state \probbar \messg)
\cdot \EU(\state,\messg,\Sstrat)\,.
\end{align*}
The discrete noise-perturbed replicator dynamics on behavioral
strategies proposed by \citet{Correia2013:The-Bivalent-Tr} is:
\begin{align*}
  \Sstrat'(\messg \probbar \state_i) & = \sum_{j} C_{ji} \cdot
  \frac{\Sstrat(\messg \probbar \state_j) \cdot
    \EU(\messg,\state_j,\Rstrat)} {\Phi(\state_j,\Rstrat)} \\
    \Rstrat'(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
  \frac{\Rstrat(\state_j \probbar \messg) \cdot
    \EU(\state_j,\messg,\Sstrat)} {\Phi(\messg,\Sstrat)}  \,.
\end{align*}
Notice that this update is also sequential: first we calculate an
update according to the standard replicator dynamics (in behavioral
strategies), then we compute the perturbation from perceptual noise.

### 2014-08-26 paper/03_simulations.tex
data exploration MF

We perform a parameter sweep with size of the state space $n \in \{10, 50, 100, 200\}$ and impairment $\impairment \in \{0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4\}$.
For each parameter combination, we perform $100$ trials and record the values of each of the metrics introduced in Section~\ref{sec:metrics} after convergence, as well as two additional indicators: number of iterations taken until convergence, and the normalized expected utility.
Expected utility is normalized by dividing by the expected utility of the theoretical optimum without impairment, \emph{i.e.}~a Voronoi tessalation of the state space.

>>>>

We perform a parameter sweep with size of the state space $n \in \{10,
50, 90, 130\}$ and impairment $\impairment \in \{0, 0.1, 0.2, 0.3,
0.4, 0.5\}$.
For each parameter combination, we perform $100$ trials and record the values of each of the metrics introduced in Section~\ref{sec:metrics} after convergence, as well as two additional indicators: number of iterations taken until convergence, and the normalized expected utility.
Expected utility is normalized by dividing by the expected utility of the theoretical optimum without impairment, \emph{i.e.}~a Voronoi tessalation of the state space.
